How to obtain the redshift distribution from probabilistic redshift estimates

### 2019-06-21 research/paper/draft.tex
responding to feedback from Hogg throughout

\usepackage{soul}
%\usepackage[title]{appendix}

\newcommand{\todo}[3]{{\color{#2}\emph{#1}: #3}}
\newcommand{\aim}[1]{\todo{AIM}{red}{#1}}
\newcommand{\myemail}{aimalz@nyu.edu}
\newcommand{\mathul}[1]{\underline{#1}}
\newcommand{\Sect}[1]{Section~\ref{#1}}
\newcommand{\Eq}[1]{Equation~\ref{#1}}
\newcommand{\Fig}[1]{Figure~\ref{#1}}
\newcommand{\Tab}[1]{Table~\ref{#1}}
\newcommand{\project}[1]{{\textsc{#1}}}
\newcommand{\lsst}{\project{LSST}}
\newcommand{\desc}{\lsst-\project{DESC}}
\newcommand{\sdss}{\project{SDSS}}
\newcommand{\boss}{\project{BOSS}}
\newcommand{\des}{\project{DES}}
\newcommand{\Chippr}{\project{CHIPPR}}
\newcommand{\repo}[1]{{\texttt{#1}}~}
\newcommand{\qp}{\repo{qp}}
\newcommand{\chippr}{\repo{chippr}}
\newcommand{\cosmolike}{\repo{CosmoLike}}
\newcommand{\emcee}{\repo{emcee}}
\newcommand{\github}{\href{https://github.com}{GitHub}}
\newcommand{\python}{\textit{Python}}
\newcommand{\nz}{$n(z)$}
\newcommand{\data}{\ensuremath{\vec{d}}}
% could change to bold
\newcommand{\like}{\mathscr{L}}
\newcommand{\pr}[1]{\ensuremath{\mathrm{p}(#1)}}
% could change to Prob or Pr 
\newcommand{\expect}[1]{\left<#1\right>}
\newcommand{\normal}[2]{\mathcal{N} (#1, #2)}
\newcommand{\gvn}{\mid}% could use | or \vert
\newcommand{\integral}[2]{\ensuremath{\int #1 \mathrm{d} #2}}
\newcommand{\sz}{spec-$z$~}
\newcommand{\Sz}{Spec-$z$~}
\newcommand{\pz}{photo-$z$~}
\newcommand{\Pz}{Photo-$z$~}
\newcommand{\pzpdf}{\pz PDF}
% could change to posterior
\newcommand{\Pzpdf}{\Pz PDF}
% could change to posterior
\newcommand{\zpdf}{redshift posterior}
% could change to implicit posterior
\newcommand{\pzip}{\pz implicit posterior}
\newcommand{\Nz}{$N(z)$~}
\newcommand{\stack}{$\hat{N}(z)$~}
\newcommand{\ntot}{\ensuremath{N_{\mathrm{tot}}}}
\newcommand{\bvec}[1]{\ensuremath{\boldsymbol{#1}}}
% could change to \vec
\newcommand{\ndphi}{\bvec{\phi}}
\newcommand{\mmle}{marginalized maximum likelihood estimate}
% really marginalized maximum a posteriori estimate, may still change this

@@

\begin{abstract}
A trustworthy estimate of the redshift distribution \nz\ is crucial for using weak gravitational lensing and large-scale structure of galaxy catalogs to study cosmology.
Spectroscopic redshifts for the dim and numerous galaxies of weak-lensing surveys are expected to be unavailable, making photometric redshifts (photo-$z$s) the next-best alternative.
The nontrivial systematics affecting photo-$z$ estimation have motivated the weak-lensing community to favor photo-$z$ probability density functions (PDFs) as a more comprehensive alternative to photo-$z$ point estimates.
However, analytic methods for utilizing these new data products in cosmological inference are still evolving.
The ubiquitous methodology known as stacking produces a systematically biased estimator of $n(z)$ that worsens with decreasing signal-to-noise, the very regime where photo-$z$ PDFs are most necessary.
We introduce a mathematically rigorous probabilistic graphical model (PGM) of $n(z)$ for hierarchical inference, which is provably the only self-consistent way to combine photo-$z$ PDFs to produce an estimator of $n(z)$.
Our Cosmological Hierarchical Inference with Probabilistic Photometric Redshifts (\Chippr) model yields a more accurate characterization of $n(z)$ by correctly propagating the redshift uncertainty information beyond the best-fit estimator produced by traditional procedures.
The CHIPPR approach is applicable to any one-point statistic of any random variable.
We conclude by propagating these effects to constraints in the space of cosmological parameters, .
\aim{be clear that rigor comes at a cost, chippr makes explicit the requirements that the pzpdfs be posteriors under a known prior (which would be required by any justifiable method), be honest that we're not ready for prime time}
\end{abstract}

@@

Photometric redshift (\pz) estimation has been a staple of studies of galaxy evolution, large-scale structure, and cosmology since its conception decades ago \citep{baum_photoelectric_1962}.  
An extremely coarse spectrum in the form of photometry in a handful of broadband filters can be an effective substitute for the time- and photon-intensive process of obtaining a spectroscopic redshift (\sz), a procedure that may only be applied to relatively bright galaxies.  
Once the photometric colors are calibrated against either a library of spectral energy distribution (SED) templates or a data set of spectra for galaxies with known redshifts, a correspondence between photometric colors and redshifts may be constructed, forming a trustworthy basis for \pz\ estimation or testing.

Calculations of correlation functions of cosmic shears and galaxy positions that constrain the cosmological parameters require large numbers of high-confidence redshifts of surveyed galaxies.  
Many more \pz s may be obtained in the time it would take to observe a smaller number of \sz s, and \pz s may be measured for galaxies too dim for accurate \sz\ confirmation, permitting the compilation of large catalogs of galaxy redshifts spanning a broad range of redshifts and luminosities.  
\Pz s have thus enabled the era of precision cosmology, heralded by weak gravitational lensing tomography and baryon acoustic oscillation peak measurements.

However, \pz s are susceptible to inaccuracy and imprecision via a number of effects, particularly their inherent noisiness due to the coarseness of photometric filters, catastrophic errors in which galaxies of one type at one redshift are mistaken for galaxies of another type at a different redshift, and systematics introduced by observational techniques, data reduction processes, and training or template set limitations.  
Figure~\ref{fig:pedagogical_scatter} illustrates the relationship between photometry and redshift generically by showing ``data'' in one dimension for visualization purposes, suggesting that a special nonlinear projection of the photometry could more-or-less yield a one-to-one relationship with true redshifts.
This plot looks very much like the traditional $z_{\mathrm{spec}}$ versus $z_{\mathrm{phot}}$ plots of \pz\ point estimates and was indeed adapted from one such plot in \citet{jain_whole_2015}, because \pz\ point estimates effectively are a special nonlinear projection of the data.

@@

There are several varieties of generally non-Gaussian deviation from a perfect correspondence between redshift and data in Figure~\ref{fig:pedagogical_scatter}, represented by a $y = x$ diagonal line.
There is scatter about the diagonal due to the coarseness of the photometric filters, with larger scatter perpendicular to the diagonal due to the specific wavelengths where highly identifiable spectral features pass between the filters, as well as higher scatter at high redshifts due to larger photometric errors on fainter galaxies.
There are populations of outliers, far from the diagonal, comprised of galaxies for which the redshift estimate is catastrophically distinct from the true redshift, showing that outliers are not uniformly distributed nor restricted to long tails away from a Gaussian scatter.
And, though hardly perceptible in the plot, there is a systematic bias, wherein the average of the points would not lie on the diagonal but would be offset by a small amount, suggested by the trend of points to lie below the diagonal at high redshift.
Based on the goals of a photometric galaxy survey, limits can be placed on the tolerance to these effects.
For example, the Science Requirements Document \citep{mandelbaum_weak_2017} states \lsst's requirements for the main cosmological sample, reproduced in Table~\ref{tab:lsstsrd}.

@@

Once propagated through the calculations of correlation functions of cosmic shear and galaxy positions, \pz\ errors are not insignificant contributors to the total uncertainties reported on cosmological parameters.
As progress has been made on the influence of other sources of systematic error, the uncertainties associated with \pz s have come to dominate the error budget of cosmological parameter estimates made by current surveys such as \des\ \citep{hoyle_dark_2017}, \project{HSC} \citep{tanaka_photometric_2018}, and \project{KiDS} \citep{hildebrandt_kids-450:_2017}.

Much effort has been dedicated to improving \pz s, though they are still most commonly obtained by a maximum likelihood estimator (MLE) based on libraries of galaxy SED templates, with conservative approaches to error estimation.
This approach tends to lead to catastrophic outliers like the horizontally oriented population of \Fig{fig:pedagogical_scatter}, due to the presence of galaxies whose SEDs are not represented by the template library.
Data-driven approaches tend to result in catastrophic outliers like the vertically oriented population of \Fig{fig:pedagogical_scatter}, due to training sets that are incomplete in redshift coverage.
Recent advances have focused on identifying and removing catastrophic outliers when using \pz s for 
inference \citep{gorecki_new_2014}.  
The approaches of using a training set versus a template library are related to one another by \citet{budavari_unified_2009}.
Sophisticated Bayesian techniques and cutting-edge machine learning methods have been employed to improve precision \citep{carliles_random_2010} and accuracy \citep{sadeh_annz2:_2016}.

An alternative to point estimation of \pz s is redshift probability distribution function (PDF) estimation that reports the probability $\pr{z}$ over all possible redshifts for every galaxy rather than an MLE (with or without presumed Gaussian error bars) \citep{koo_photometric_1999}.  
This option is favorable because it contains more potentially useful information about the uncertainty on each galaxy's redshift, incorporating notions of precision, accuracy, and systematic errors.
However, denoting \pzpdf s as ``$\pr{z}$'' is an abuse of notation, as it does not adequately convey what information is being used to constrain the redshift $z$; \pzpdf s are \textit{posterior} PDFs, conditioned on the photometric data and prior knowledge.
In terms of \Fig{fig:pedagogical_scatter}, \pzpdf s are horizontal cuts, probabilities of redshift conditioned on a specific value of data, i.e. posteriors, which constrain redshifts, whereas vertical cuts through this space are probabilities of data conditioned on a specific redshift, i.e. likelihoods, from which data is actually drawn.

\Pzpdf s have been produced by completed surveys \citep{hildebrandt_cfhtlens:_2012, sheldon_photometric_2012} and will be produced by ongoing and upcoming surveys \citep{abell_lsst_2009, carrasco_kind_exhausting_2014, bonnett_redshift_2016, masters_mapping_2015}.  
\Pzpdf s are not without their own weaknesses, however, including the resources necessary to calculate and record them for large galaxy surveys \citep{carrasco_kind_sparse_2014} and the divergent results of each method used to derive them \citep{hildebrandt_phat:_2010, dahlen_critical_2013, sanchez_clustering_2013, bonnett_redshift_2016, tanaka_photometric_2018}.  
The most important of these issues, however, is that use of them in the literature is inconsistent at best and incorrect at worst.  
The most common application of \pzpdf s is their use in estimating \Nz, the distribution of redshifts of a sample of galaxies, a quantity essential to the calculation of the power spectra of weak gravitational lensing and large-scale structure that are used to constrain the parameters of cosmological models \citet{mandelbaum_precision_2008, sheldon_photometric_2012, bonnett_redshift_2016}.
%\aim{Provide equations/citations for the use of \Nz\ in cosmology.}

If the true redshifts $\{z_{i}^{\dagger}\}$ of galaxies $i$ were known, then the redshift PDFs would be delta functions $\{\delta(z, z_{i}^{\dagger})\}$ centered at the true redshift, and the redshift distribution could be effectively approximated by a histogram or other interpolation of the delta functions $\{\delta(z, z_{i}^{\dagger})\}$.
When \pzpdf s are available instead of true redshifts, the simplest approach reduces them to point estimates $\{\hat{z}_{i}\}$ of redshift by using $\delta(z, \hat{z}_{i})$ in place of $\delta(z, z_{i}^{\dagger})$.
Though it is most common for $\hat{z}_{i}$ to be the mode (maximum) of the \pzpdf, there are other, more principled point estimate reduction procedures \citep{tanaka_photometric_2018}.

To circumvent the reduction of uncertainty information due to point estimation, however, it has grown to be more common to calculate the mathematically inconsistent but conceptually simple \textit{stacked estimator} of the redshift density function \citep{lima_estimating_2008}, 
\begin{align}
\label{eqn:stack}
\hat{n}(z) &= \frac{1}{\ntot} \sum_{i = 0}^{\ntot} \pr{z}_{i}
\end{align}
for a sample of $\ntot$ galaxies $i$, or, equivalently, the redshift distribution function $\hat{N}(z) = \ntot \hat{n}(z)$.
Much of this thesis centers around the problems with this logically invalid yet pervasive quantity.

\aim{TODO: Pull in content from global intro of thesis.}

Though their potential to improve estimates of physical parameters is tremendous, \pzpdf s have been applied only to a limited extent.  
They have been used to form selection criteria of samples from galaxy surveys without propagation 
through the calculations of physical parameters \citep{van_breukelen_reliable_2009, viironen_high_2015}.  
Probability cuts on Bayesian quantities are not uncommon \citep{leung_bayesian_2017, dipompeo_quasar_2015}, but that procedure does not fully take advantage of all information contained in a probability distribution for parameter inference.

Despite the growing prevalence of \pzpdf\ production, no implementation of inference using \pzpdf s has yet been presented with a mathematically consistent methodology.  
We present and validate a hierarchical Bayesian technique for the use of \pzpdf\ in inference of arbitrary statistics relevant to cosmology, large-scale structure, and galaxy evolution.  
For simplicity, we consider only one-point statistics, though future work will extend this methodology to higher-order statistics.

The \textit{redshift distribution function \Nz}, or, almost interchangably, its normalized cousin the \textit{redshift density function \nz}, serves as an ideal statistic upon which to demonstrate this novel approach.  
\Nz\ is necessary for calculations of two-point correlation functions of weak gravitational lensing and counting statistics that are used to probe dark energy \citep{masters_mapping_2015}.  
The observed \Nz\ also been used to validate survey selection functions usnfenjed in generation of realistic, multi-purpose mock catalogs \citep{norberg_2df_2002}.  
Additionally, \Nz\ has been the subject of inference using \pzpdf s before \citep{sheldon_photometric_2012, hildebrandt_cfhtlens:_2012, kelly_weighing_2014, benjamin_cfhtlens_2013, bonnett_using_2015, viironen_high_2015, asorey_galaxy_2016, leistedt_hierarchical_2016}, so comparisons to the literature may easily be made.

The prevailing way to estimate \nz\ for a sample of galaxies $j$ is to ``stack'' \pzpdf s according to Equation~\ref{eqn:stack}.
The stacked estimator of the redshift density function is effectively an average of the \pzpdf\ catalog and, as we show in this paper, not in general mathematically valid.
Stacking is nonetheless considered the preferred method for obtaining \Nz\ from a \pzpdf\ catalog \citep{sheldon_photometric_2012, kelly_weighing_2014, benjamin_cfhtlens_2013, bonnett_using_2015, viironen_high_2015, asorey_galaxy_2016}.  
However, it must be noted here that Equation~\ref{eqn:stack} 
is not in general mathematically valid.  
(See \citet{hogg_data_2012} for a complete discussion.)

%\aim{Motivate how well we need to know \Nz\ and why it matters.
%	New paragraph here: Say what precision is needed for \Nz\ for future weak lensing surveys. 
%	Say what precision the mass function is needed (in, say cluster studies) for precision cosmology.}

We aim to develop a clear methodology guiding the use of \pzpdf s in inference so they may be utilized effectively by the cosmology community.
Though others have approached the problem before \citep{leistedt_hierarchical_2016}, the method presented here differs in that it makes use of any existing catalog of \pzpdf s, rather than requiring a simultaneous derivation of the \pzpdf s and the redshift distribution, making it preferable to ongoing surveys that may be resistant to restructuring their analysis pipelines.

@@

The redshift distribution function \nz\ is the number of galaxies per unit redshift, effectively defining the evolution in the number of galaxies \citep{menard_clustering-based_2013}.  
% \aim{also selection function}
In the following sections, we present and compare methods for estimating \nz\ from \pzpdf s.  
\Sect{sec:prob} contains the mathematical derivation of a probabilistic model for $n(z)$ dependent on photo-$z$ probability distribution functions, and \Sect{sec:sheldon} contrasts the probabilistic model with alternative methods.

@@

\aim{TODO: fix implicit posterior vs. implicit-weighted posterior}

@@

\begin{figure}
	\begin{center}
		\includegraphics[width=0.15\textwidth]{figures/chippr/pgm.png}
		\caption{The directed acyclic graph of the CHIPPR model, where circles indicate random variables and arrows indicate causal relationships.
			The redshift distribution function parameterized by hyperparameters $\ndphi$ exists independent of the survey of $J$ galaxies, indicated as a box.  
			The redshifts $\{z_{j}\}$ of all galaxies in the survey are latent variables independently drawn from the redshift distribution, which is a function of $\ndphi$. 
			The photometric data $\data_{j}$ for each galaxy is drawn from a function of its redshift $z_{j}$ and observed, indicated by a shaded circle.}
		\label{fig:pgm}
	\end{center}
\end{figure}

The problem facing cosmologists is to determine the true value of $\ndphi$ from observing the photometry $\{\data_{j}\}$ of a large sample of $J$ galaxies $j$.
To self-consistently propagate the uncertainty in the redshifts, however, it is more appropriate to estimate the posterior $\pr{\ndphi \gvn \{\data_{j}\}}$ over all possible values of $\ndphi$ conditioned on all the observed data $\{\data_{j}\}$ available in a generic catalog.
In order to use the DAG of \Fig{fig:pgm} to derive an expression for $\pr{\ndphi \gvn \{\data_{j}\}}$ in terms of \pzpdf s, we must introduce two more concepts, confusingly named the implicit prior and the prior probability density.

>>>>

\usepackage{soul}
\usepackage[title]{appendix}

\newcommand{\todo}[3]{{\color{#2}\emph{#1}: #3}}
\newcommand{\aim}[1]{\todo{AIM}{red}{#1}}
\newcommand{\dwh}[1]{\todo{Attn. Hogg:}{blue}{#1}}
\newcommand{\myemail}{aimalz@nyu.edu}
\newcommand{\mathul}[1]{\underline{#1}}
\newcommand{\Sect}[1]{Section~\ref{#1}}
\newcommand{\Eq}[1]{Equation~\ref{#1}}
\newcommand{\Fig}[1]{Figure~\ref{#1}}
\newcommand{\Tab}[1]{Table~\ref{#1}}
\newcommand{\project}[1]{{\textsc{#1}}}
\newcommand{\lsst}{\project{LSST}}
\newcommand{\desc}{\lsst-\project{DESC}}
\newcommand{\sdss}{\project{SDSS}}
\newcommand{\boss}{\project{BOSS}}
\newcommand{\des}{\project{DES}}
\newcommand{\Chippr}{\project{CHIPPR}}
\newcommand{\repo}[1]{{\texttt{#1}}~}
\newcommand{\qp}{\repo{qp}}
\newcommand{\chippr}{\repo{chippr}}
\newcommand{\cosmolike}{\repo{CosmoLike}}
\newcommand{\emcee}{\repo{emcee}}
\newcommand{\github}{\href{https://github.com}{GitHub}}
\newcommand{\python}{\textit{Python}}
\newcommand{\data}{\ensuremath{\vec{d}}}
% could change to bold
\newcommand{\like}{\mathscr{L}}
\newcommand{\pr}[1]{\ensuremath{\mathrm{p}(#1)}}
% could change to Prob or Pr 
\newcommand{\expect}[1]{\left<#1\right>}
\newcommand{\normal}[2]{\mathcal{N} (#1, #2)}
\newcommand{\gvn}{\mid}% could use | or \vert
\newcommand{\integral}[2]{\ensuremath{\int #1 \mathrm{d} #2}}
\newcommand{\sz}{spec-$z$}
\newcommand{\Sz}{Spec-$z$}
\newcommand{\pz}{photo-$z$}
\newcommand{\Pz}{Photo-$z$}
\newcommand{\pzpdf}{\pz\ PDF}
% could change to posterior
\newcommand{\Pzpdf}{\Pz\ PDF}
% could change to posterior
\newcommand{\zpdf}{redshift posterior}
% could change to implicit posterior
\newcommand{\pzip}{\pz implicit posterior}
\newcommand{\nz}{$n(z)$}
\newcommand{\Nz}{$N(z)$}
\newcommand{\stack}{$\hat{n}(z)$}
\newcommand{\ntot}{\ensuremath{N_{\mathrm{tot}}}}
\newcommand{\bvec}[1]{\ensuremath{\boldsymbol{#1}}}
% could change to \vec
\newcommand{\ndphi}{\bvec{\phi}}
\newcommand{\mmle}{marginalized maximum likelihood estimate}
% really marginalized maximum a posteriori estimate, may still change this

@@

\begin{abstract}
A trustworthy estimate of the redshift distribution \nz\ is crucial for using weak gravitational lensing and large-scale structure of galaxy catalogs to study cosmology.
Spectroscopic redshifts for the dim and numerous galaxies of weak-lensing surveys are expected to be unavailable, making photometric redshifts (photo-$z$s) the next-best alternative.
The nontrivial systematics affecting photo-$z$ estimation have motivated the weak-lensing community to favor photo-$z$ posterior probability density functions (PDFs) as a more comprehensive alternative to photo-$z$ point estimates.
However, analytic methods for utilizing these new data products in cosmological inference are still evolving.
The ubiquitous methodology known as stacking produces a systematically biased estimator of $n(z)$ that worsens with decreasing signal-to-noise, the very regime where photo-$z$ PDFs are most necessary.
We introduce a statistically rigorous probabilistic graphical model (PGM) of $n(z)$ for hierarchical inference, which is provably the only self-consistent way to combine photo-$z$ PDFs to produce an estimator of $n(z)$.
Our Cosmological Hierarchical Inference with Probabilistic Photometric Redshifts (\Chippr) model yields a more accurate characterization of $n(z)$ by correctly propagating the redshift uncertainty information beyond the best-fit estimator of $n(z)$ produced by traditional procedures.
Using the publicly available \texttt{chippr} prototype code, we propagate these effects to constraints in the space of cosmological parameters, noting, however, that the mathematically justifiable approach incurs computational expense.
The CHIPPR approach is applicable to any one-point statistic of any random variable, provided the prior probability distribution used to produce the posteriors is known.
\end{abstract}

@@

Photometric redshift (\pz) estimation has been a staple of studies of galaxy evolution, large-scale structure, and cosmology since its conception half a century ago \citep{baum_photoelectric_1962}.  
An extremely coarse spectrum in the form of photometry in a handful of broadband filters can be an effective substitute for the time- and photon-intensive process of obtaining a spectroscopic redshift (\sz), a procedure that may only be applied to relatively bright galaxies.  
Once the photometric colors are calibrated against either a library of spectral energy distribution (SED) templates or a data set of spectra for galaxies with known redshifts, a correspondence between photometric colors and redshifts may be constructed, forming a trustworthy basis for \pz\ estimation or testing.

Calculations of correlation functions of cosmic shears and galaxy positions that constrain the cosmological parameters require large numbers of high-confidence redshifts of surveyed galaxies.  
Many more \pz s may be obtained in the time it would take to observe a smaller number of \sz s, and \pz s may be measured for galaxies too dim for accurate \sz\ confirmation, permitting the compilation of large catalogs of galaxies spanning a broad range of redshifts and luminosities.  
\Pz s have thus enabled the era of precision cosmology, heralded by weak gravitational lensing tomography and baryon acoustic oscillation peak measurements.

However, \pz s are susceptible to inaccuracy and imprecision via a number of effects, particularly their inherent noisiness due to the coarseness of photometric filters, catastrophic errors in which galaxies of one SED at one redshift are mistaken for galaxies of another SED at a different redshift, and systematics introduced by observational techniques, data reduction processes, and training or template set limitations.  
Figure~\ref{fig:pedagogical_scatter} illustrates the relationship between photometry and redshift generically by showing ``data'' in one dimension for visualization purposes, suggesting that a special nonlinear projection of the photometry could more-or-less yield a one-to-one relationship with true redshifts.
The similarity to traditional $z_{\mathrm{spec}}$ versus $z_{\mathrm{phot}}$ plots of \pz\ point estimates is no coincidence, as \pz\ point estimates effectively are indeed one such special nonlinear projection of the data.

@@

There are several varieties of generally non-Gaussian deviation from a perfect correspondence between redshift and data in Figure~\ref{fig:pedagogical_scatter}, represented by a $y = x$ diagonal line.
There is scatter about the diagonal due to the coarseness of the photometric filters, with larger scatter perpendicular to the diagonal due to the specific wavelengths where highly identifiable spectral features pass between the filters, as well as higher scatter at high redshifts due to larger photometric errors on fainter galaxies.
There are populations of outliers, far from the diagonal, comprised of galaxies for which the redshift estimate is catastrophically distinct from the true redshift, showing that outliers are not uniformly distributed nor restricted to long tails away from a Gaussian scatter.
And, though hardly perceptible in the plot, there is a systematic bias, wherein the average of the points would not lie on the diagonal but would be offset by a small bias, suggested by the trend of high redshift points to lie below the diagonal.

Once propagated through the calculations of correlation functions of cosmic shear and galaxy positions, \pz\ errors are not insignificant contributors to the total uncertainties reported on cosmological parameters.
As progress has been made on the influence of other sources of systematic error, the uncertainties associated with \pz s have come to dominate the error budget of cosmological parameter estimates made by current surveys such as \des\ \citep{hoyle_dark_2017}, \project{HSC} \citep{tanaka_photometric_2018}, and \project{KiDS} \citep{hildebrandt_kids-450:_2017}.
Based on the goals of a photometric galaxy survey, limits can be placed on the tolerance to these effects.
For example, the Science Requirements Document \citep{mandelbaum_weak_2017} states \lsst's requirements for the main cosmological sample, reproduced in Table~\ref{tab:lsstsrd}.

@@

Much effort has been dedicated to improving \pz s, though they are still most commonly obtained by a maximum likelihood estimator (MLE) based on libraries of galaxy SED templates, with conservative approaches to error estimation.
This approach tends to lead to catastrophic outliers like the horizontally oriented population of \Fig{fig:pedagogical_scatter}, due to the presence of galaxies whose SEDs are not represented by the template library.
Data-driven approaches tend to result in catastrophic outliers like the vertically oriented population of \Fig{fig:pedagogical_scatter}, due to training sets that are incomplete in redshift coverage.
The approaches of using a training set versus a template library are related to one another by \citet{budavari_unified_2009}.
Sophisticated Bayesian techniques and machine learning methods have been employed to improve precision \citep{carliles_random_2010} and accuracy \citep{sadeh_annz2:_2016}, while other advances have focused on identifying and removing catastrophic outliers when using \pz s for inference \citep{gorecki_new_2014}.

An alternative to point estimation of \pz s is redshift probability distribution function (PDF) estimation that reports the probability of all possible redshifts for every galaxy, commonly written as $\pr{z}$, rather than an MLE (with or without presumed Gaussian error bars) \citep{koo_photometric_1999}.  
This option is favorable because it contains more potentially useful information about the uncertainty on each galaxy's redshift, incorporating our understanding of precision, accuracy, and systematic error.
However, denoting \pzpdf s as ``$\pr{z}$'' is an abuse of notation, as it does not adequately convey what information is being used to constrain the redshift $z$; \pzpdf s are \textit{posterior} PDFs, conditioned on the photometric data and prior knowledge.
In terms of \Fig{fig:pedagogical_scatter}, \pzpdf s are horizontal cuts, probabilities of redshift conditioned on a specific value of data, i.e. posteriors, which constrain redshifts, whereas vertical cuts through this space are probabilities of data conditioned on a specific redshift, i.e. likelihoods, from which data is actually drawn.

\Pzpdf s have been produced by completed surveys \citep{hildebrandt_cfhtlens:_2012, sheldon_photometric_2012} and will be produced by ongoing and upcoming surveys \citep{abell_lsst_2009, carrasco_kind_exhausting_2014, bonnett_redshift_2016, masters_mapping_2015}.  
\Pzpdf s are not without their own shortcomings, however, including the resources necessary to calculate and record them for large galaxy surveys \citep{carrasco_kind_sparse_2014, malz_approximating_2018} and the divergent results of each method used to derive them \citep[ and reviewed in Schmidt, Malz \& Soo, et al. (in prep)]{hildebrandt_phat:_2010, dahlen_critical_2013, sanchez_clustering_2013, bonnett_redshift_2016, tanaka_photometric_2018}.  
The most concerning weakness of \pzpdf s, however, is their usage in the literature, which is at best inconsistent and at worst incorrect.  
The most prevalent application of \pzpdf s is their use in estimating the \textit{redshift distribution function \Nz}, or, interchangably, its normalized cousin the \textit{redshift density function \nz}, a quantity essential to the calculation of the power spectra of weak gravitational lensing and large-scale structure that are used to constrain the parameters of cosmological models \citep{mandelbaum_precision_2008, hildebrandt_cfhtlens:_2012, sheldon_photometric_2012, benjamin_cfhtlens_2013, kelly_weighing_2014, bonnett_using_2015,  masters_mapping_2015, viironen_high_2015, asorey_galaxy_2016, bonnett_redshift_2016, leistedt_hierarchical_2016}, as well as to validate survey selection functions used in generation of realistic, multi-purpose mock catalogs \citep{norberg_2df_2002}.

\aim{TODO:
	Motivate how well we need to know \Nz\ and why it matters.
	Provide equations/citations for the use of \Nz\ in cosmology.
	New paragraph here: Say what precision is needed for \Nz\ for future weak lensing surveys. 
	Say what precision the mass function is needed (in, say cluster studies) for precision cosmology.}

If the true redshifts $\{z_{j}^{\dagger}\}$ of galaxies $j$ are known, then the redshift PDFs are delta functions $\{\delta(z, z_{j}^{\dagger})\}$ centered at the true redshift, and the redshift distribution is effectively approximated by a histogram or other interpolation of the delta functions $\{\delta(z, z_{j}^{\dagger})\}$.
When \pzpdf s are available instead of true redshifts, the simplest approach reduces them to point estimates $\{\hat{z}_{i}\}$ of redshift by using $\delta(z, \hat{z}_{j})$ in place of $\delta(z, z_{j}^{\dagger})$.
Though it is most common for $\hat{z}_{j}$ to be the maximum or \textit{mode} of the \pzpdf, there are other, more principled point estimate reduction procedures \citep{tanaka_photometric_2018}.

However, any procedure that reduces \pzpdf s to point estimates loses information about the uncertainty on redshift.
To circumvent this loss, it has grown to be more common to calculate the conceptually simple but mathematically inconsistent \citep{hogg_data_2012} \textit{stacked estimator} $\hat{n}(z)$ of the redshift density function \citep{lima_estimating_2008}
\begin{align}
\label{eqn:stack}
\hat{n}(z) &= \frac{1}{J} \sum_{j = 0}^{J} \pr{z}_{j}
\end{align}
for a sample of $J$ galaxies $j$, or, equivalently, the redshift distribution function $\hat{N}(z) = J \hat{n}(z)$, by effectively averaging the \pzpdf s.

%Though their potential to improve estimates of physical parameters is tremendous, \pzpdf s have been applied only to a limited extent.  
%They have been used to form selection criteria of samples from galaxy surveys without propagation 
%through the calculations of physical parameters \citep{van_breukelen_reliable_2009, viironen_high_2015}.  
%Probability cuts on Bayesian quantities are not uncommon \citep{leung_bayesian_2017, dipompeo_quasar_2015}, but that procedure does not fully take advantage of all information contained in a probability distribution for parameter inference.

Despite the growing prevalence of \pzpdf\ production, no implementation of inference using \pzpdf s has yet been presented with a mathematically consistent methodology.  
This paper challenges the logically invalid yet pervasive analysis procedure of stacking \pzpdf s by presenting and validating a hierarchical Bayesian technique for the use of \pzpdf s\ in the inference of \nz, yielding a method applicable to arbitrary one-point statistics relevant to cosmology, large-scale structure, and galaxy evolution; future work will extend this methodology to higher-order statistics.
We aim to develop a clear methodology guiding the use of \pzpdf s in inference so they may be utilized effectively by the cosmology community.
Though others have approached the problem before \citep{leistedt_hierarchical_2016}, the method presented here differs in that it makes use of any existing catalog of \pzpdf s, rather than requiring a simultaneous derivation of the \pzpdf s and the redshift distribution, making it preferable to ongoing surveys for which there may be inertia preventing a complete restructuring of the analysis pipeline.

@@

The redshift distribution function \nz\ is the number of galaxies per unit redshift, effectively defining the evolution in the number of galaxies convolved with the selection function of the sample \citep{menard_clustering-based_2013}.  
In the following sections, we present and compare methods for estimating \nz\ from \pzpdf s.  
\Sect{sec:prob} contains the mathematical derivation of a probabilistic model for $n(z)$ dependent on photo-$z$ probability distribution functions, and \Sect{sec:sheldon} contrasts the probabilistic model with alternative methods.

@@

@@

\begin{figure}
	\begin{center}
		\includegraphics[width=0.15\textwidth]{figures/chippr/pgm.png}
		\caption{The directed acyclic graph of the CHIPPR model, where circles indicate random variables and arrows indicate causal relationships.
			The redshift distribution \nz\ parameterized by hyperparameters $\ndphi$ exists independent of the survey of $J$ galaxies, indicated as a box.  
			The redshifts $\{z_{j}\}$ of all galaxies in the survey are latent variables independently drawn from the redshift distribution, which is a function of $\ndphi$. 
			The photometric data $\data_{j}$ for each galaxy is drawn from a function of its redshift $z_{j}$ and observed, indicated by a shaded circle.}
		\label{fig:pgm}
	\end{center}
\end{figure}

The problem facing cosmologists is to determine the true value of $\ndphi$ from observing the photometry $\{\data_{j}\}$ of a large sample of $J$ galaxies $j$.
To self-consistently propagate the uncertainty in the inference of redshift, however, it is more appropriate to estimate the posterior $\pr{\ndphi \gvn \{\data_{j}\}}$ over all possible values of $\ndphi$ conditioned on all the observed data $\{\data_{j}\}$ available in a generic catalog.
In order to use the DAG of \Fig{fig:pgm} to derive an expression for $\pr{\ndphi \gvn \{\data_{j}\}}$ in terms of \pzpdf s, we must introduce two more concepts, confusingly named the \textit{implicit prior} and the \textit{prior probability density (prior PDF)}, elaborated upon below.

\dwh{When we constrain the redshift of a galaxy using its observed photometric data $\data_{j}$, we are effectively estimating a posterior $\pr{z \gvn \data_{j}}$, the probability of an unknown quantity conditioned on the quantity we have in hand, i.e the data.
However, to do this, we must have a model for the general relationship between redshifts and photometry, whether empirical, as is the case for machine learning \pzpdf\ methods, or analytic, as is the case for template-based \pzpdf\ methods.
Such a relationship $\pr{\data, z}$ is defined in the multidimensional space of probability density over redshift and photometry.
If we were to marginalize over the photometry in $\pr{\data, z}$, we would obtain a one-dimensional PDF $\pr{z \gvn \ndphi^{*}}$ over redshift, which can by definition be parameterized by the same functional form as \nz, for some $\ndphi^{*}$ specific to the estimation procedure that may or may not bear any relation to the hyperparameters $\ndphi^{\dagger}$ of the true \nz.
Rather, $\ndphi^{*}$ is a consequence of generative model for how photometry results from redshift, including the influence of intrinsic galaxy spectra and instrumental effects. 
We call $\pr{z \gvn \ndphi^{*}}$ the \textit{implicit prior}, as it is rarely explicitly known nor chosen by the researcher\footnote{For template-based methods, the implicit prior is often an explicitly known input to the algorithm, engineered as an initial guess for the true $\ndphi$, with an aim for a realistic choice guided by an earlier spectroscopic survey.  
(See \citet{benitez_bayesian_2000} for more detail.)
It may thus be more appropriate to call it an \textit{interim prior}, but we will use the former term throughout this paper for generality.}
Because the implicit prior is unavoidable and almost inherently not uninformative, the \pzpdf s reported by any method must be \textit{implicit posteriors} $\pr{z \gvn \data, \ndphi^{*}}$ weighted by the implicit prior.}
